Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-07-05 WSSSPE2_report/WSSSPE2_review.tex
Section 9.1: minor changes.

This group discussed five papers with a wide variety ideas of how to support
reproducibility and reuse. It focused on identifying concrete that the attendees
could work together on, which would have a positive effect on the community.

@@

In the first discussion session, this group broke up into two smaller groups
working on different topics: Reproducibility, and Reuse and Sharing. In the
second discussion session, three groups worked on creating specific pieces of
guidance.

@@

If journals do enforce greater reproducibility constraints, it is important to
lower the barriers to reviewers attempting to verifying the correctness of the
software used to generate the results. A major issue is that a lot of software
only builds on certain systems. Should journals provide more tools/support for
reviewers, and if so, what is it? An idea that came from the groups was to
define a set of support services that should be available to software paper
reviewers. Another was to provide the ability for reviewers to flag the
requirement for a `software verification,' similar to ability to flag that a
paper needs to be seen by a statistician.

Other discussion focused on ways in which researchers themselves could improve
the reproducibility of research. One way would be to establish tracks at
conferences that subject papers to reproduction, which for those that pass would
lend them an additional badge (similar to OOPSLA). Another is simply to get more
people to use your software: for instance by outreach to high school students --
can your software be used by them? Finally, there is a role for
community-curated benchmarks to validate the performance and capabilities of
tools.

@@

The second group discussed ways in which software could be more easily reused
and shared. From the perspective of both user and developer, any solutions must
be 1) easy; 2) cheap; 3) not take much time.

@@

Discovery of relevant software is still fundamental issue: we need standard
vocabularies and metadata, better tools, and approaches that are sustainable.
Publications are an easy entry point for locating suitable software, but should
the publishers lead the way, or is this the responsibility of ``the community''?
Some communities have had significant initiatives to improve software discovery,
e.g. the NIH Software Discovery Index\footnote{NIH Software Discovery Index:
\url{http://softwarediscoveryindex.org/}}. Likewise, there were examples of
journals which had made software more discoverable: ACM Transactions on Software
offer a reproducibility review; the Journal of Biostatistics has an opt-in to
provide code and a certification mark if it can be run; the Journal of Open
Research Software requires software to be deposited in suitable repositories and
referenced.

An issue around the sustainability of software catalogues is that their
usefulness often depends on the domain. For instance, in the bio-sciences there
is more homogeneous data, and more standard shred codes. In areas like ecology,
codes are very specific to the problem, meaning that the level of re-use might
be at a very general statistical level of abstraction, but then every research
use is highly customized.

Finally, it was clear from discussion that there were many ways in which
software could be made available in more reusable ways that just a tarball
sitting on a personal website. Using code repositories like
GitHub\footnote{GitHub: \url{http://github.com/}} gets you an archived, shared
platform and improve the reusability of your software incrementally, for
instance by adding a license or by archiving (with a DOI) in
Figshare\footnote{FigShare: \url{http://figshare.com/}} or
Zenodo\footnote{Zenodo: \url{http://zenodo.org/}}. Docker\footnote{Docker:
\url{htp://www.docker.com/}} might be a solution to the issue of dependencies,
to allow binaries and libraries to be bundled in a more lightweight fashion than
a virtual machine image.

>>>>

This group discussed five papers with a wide variety ideas of how to support
reproducibility and reuse. It focused on identifying concrete practices that the attendees
could work together on, which would have a positive effect on the community.

@@

In the first discussion session, this group broke up into two smaller groups
working on different topics: Reproducibility, and Reuse and Sharing. In the
second discussion session, three subgroups worked on creating specific pieces of
guidance.

@@

If journals do enforce greater reproducibility constraints, it is important to
lower the barriers to reviewers attempting to verifying the correctness of the
software used to generate the results. A major issue is that a lot of software
only builds on certain systems. Should journals provide more tools/support for
reviewers, and if so, what is it? An idea that came from the groups was to
define a set of support services that should be available to software paper
reviewers. Another was to provide the ability for reviewers to flag the
requirement for a `software verification,' similar to the ability to flag that a
paper needs to be seen by a statistician.

Other discussion focused on ways in which researchers themselves could improve
the reproducibility of research. One way would be to establish tracks at
conferences that subject papers to reproduction, which for those that pass would
lend them an additional badge (similar to OOPSLA). Another is simply to get more
people to use your software: for instance by outreach to high school students ---
can your software be used by them? Finally, there is a role for
community-curated benchmarks to validate the performance and capabilities of
tools.

@@

The second group discussed ways in which software could be more easily reused
and shared. From the perspective of both user and developer, any solutions must
be 1) easy; 2) cheap; 3) not too time-consuming.

@@

Discovery of relevant software is still fundamental issue: we need standard
vocabularies and metadata, better tools, and approaches that are sustainable.
Publications are an easy entry point for locating suitable software, but should
the publishers lead the way, or is this the responsibility of ``the community''?
Some communities have had significant initiatives to improve software discovery,
e.g., the NIH Software Discovery Index\footnote{NIH Software Discovery Index:
\url{http://softwarediscoveryindex.org/}}. Likewise, there were examples of
journals which had made software more discoverable: ACM Transactions on Software
offers a reproducibility review; the Journal of Biostatistics has an opt-in to
provide code and a certification mark if it can be run; the Journal of Open
Research Software requires software to be deposited in suitable repositories and
referenced.

An issue around the sustainability of software catalogues is that their
usefulness often depends on the domain. For instance, in the bio-sciences, there
are more homogeneous data and standard shred code. In areas like ecology,
codes are very specific to the problem, meaning that the level of re-use might
be at a very general statistical level of abstraction, but then every research
use is highly customized.

Finally, it was clear from the discussion that there were many ways in which
software could be made available in more reusable ways than just a tarball
sitting on a personal website. Using code repositories like
GitHub\footnote{GitHub: \url{http://github.com/}} gets you an archived, shared
platform and improve the reusability of your software incrementally, for
instance by adding a license or by archiving (with a DOI) in
Figshare\footnote{FigShare: \url{http://figshare.com/}} or
Zenodo\footnote{Zenodo: \url{http://zenodo.org/}}. Docker\footnote{Docker:
\url{htp://www.docker.com/}} might be a solution to the issue of dependencies,
to allow binaries and libraries to be bundled in a more lightweight fashion than
a virtual machine image.
